TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-07-30 paper/draft.tex
table caption

\newcommand{\code}[1]{\texttt{\detokenize{#1}}}
\newcommand{\apogee}{\textsl{APOGEE}}
\newcommand{\thecannon}{\textsl{The Cannon}}
\newcommand{\cannon}{\textsl{Cannon}}
\newcommand{\gaia}{\textsl{Gaia}}
\newcommand{\drtwo}{\textsl{Gaia-DR2}}
\newcommand{\zmass}{\textsl{2MASS}}
\newcommand{\sdss}{\textsl{SDSS}}
\newcommand{\kepler}{\textsl{Kepler}}
\newcommand{\lamost}{\textsl{LAMOST}}
\newcommand{\sdssiv}{\textsl{SDSS-IV}}
\newcommand{\tess}{\textsl{TESS}}
\newcommand{\msun}{M_{\odot}}
\newcommand{\teff}{T_{\mathrm{eff}}}
\newcommand{\logg}{\log g}
\newcommand{\feh}{[{\mathrm{Fe}/\mathrm{H}}]}

@@

\item \textbf{Model extraploation cuts:} Because \thecannon\ does not perform well under extrapolated regions of parameter space, we select only sources inside the range of our training sample with ASPCAP parameters $2800<T_{\rm eff}<4100$ and $-0.5<[M/H]<0.5$, and with \cannon\ parameters of $2850<T_{\rm eff}<4150$, $-0.5<\feh<0.5$, and $0<SPT<9$.

@@

\begin{deluxetable*}{cccccccc}
\tablecaption{Metallicities of \apogee\ M dwarfs with wide binaries.}
\tabletypesize{\scriptsize}
\tablehead{Mdwarf Gaia ID      & Companion Gaia ID   & Mdwarf [M/H] & Mdwarf [Fe/H] & Companion [Fe/H] & Companion source & Separation (AU)}
	\startdata
	1476768433933264128 & 1476762627137479552 & -0.18      & 0.01        & 0.10           & APOGEE Cannon M   & 16091      \\
	1476646285062821888 & 1476646250703083264 & -0.02      & 0.18        & 0.27           & APOGEE Cannon M   & 4155       \\
	4008743154907642496 & 4008743086188165248 & -0.20      & 0.02        & 0.09           & APOGEE Cannon M   & 11855      \\
	3053911383155592832 & 3053907221325174144 & -0.25      & -0.15       & -0.15          & APOGEE Cannon M   & 12884      \\
	1475867727751525632 & 1475867727751525504 & -0.25      & -0.15       & -0.13          & APOGEE Cannon M   & 5091       \\
	933708229245319552  & 933708164821511424  & -0.23      & 0.11        & 0.01           & APOGEE Cannon M   & 2245       \\
	3888520938618496768 & 3888520938618115456 & -0.16      & -0.15       & -0.14          & APOGEE Cannon M   & 1128       \\
	3888309828090932608 & 3888309832385988992 & -0.24      & -0.01       & 0.04           & APOGEE Cannon M   & 2499       \\
	374137560388179328  & 374137560388178304  & -0.34      & -0.24       & -0.34          & LAMOST FGK        & 5888       \\
	4031346880591589248 & 4031346914951327104 & -0.33      & -0.25       & -0.18          & LAMOST FGK        & 6637       \\
	3443427533502334592 & 3443427533502335744 & 0.06       & 0.20        & 0.32           & LAMOST FGK        & 3151       \\
	1605991802162861696 & 1605991767803123200 & -0.16      & -0.03       & -0.06          & LAMOST FGK        & 1043       \\
	4418850890305439872 & 4418851263967008000 & -0.18      & 0.11        & 0.27           & APOGEE ASPCAP FGK & 4920       \\
	1381455722290898176 & 1381408782592655616 & -0.13      & -0.02       & -0.08          & APOGEE ASPCAP FGK & 2401       \\
	2687603208839189376 & 2687603311918406528 & -0.12      & 0.00        & -0.03          & APOGEE ASPCAP FGK & 11201    
	\enddata
\end{deluxetable*}

@@

\begin{figure}
\begin{center}
	\includegraphics[width=\linewidth]{figures/wide_binary_met.png}
	\caption{}
\end{center}
\end{figure}

@@

As a check of internal consistency, we cross-matched our M dwarf final sample with the \citealt{Elbadry:2018b} catalog of $>$50,000 high-confidence, widely separated binaries identified by \drtwo.

As expected, the Toomre diagram in Figure \ref{fig:toomre} shows that higher metallicity sources in the sample are concentrated in low velocity space corresponding to the thin disk population, while the thick disk population contains a higher concentration of lower metallicity sources.

@@

\begin{figure*}[]
\begin{center}
	\includegraphics[width=.45\linewidth]{figures/cmd_teff_full.png}
	\includegraphics[width=.48\linewidth]{figures/cmd_teff_safe.png} \\
	\includegraphics[width=.45\linewidth]{figures/cmd_feh_full.png}
	\includegraphics[width=.48\linewidth]{figures/cmd_feh_safe.png} \\
	\includegraphics[width=.45\linewidth]{figures/cmd_spt_full.png}
	\includegraphics[width=.48\linewidth]{figures/cmd_spt_safe.png}
	\caption{Full sample of 14,827 M dwarfs colored by \cannon\ labels before selection (left) and final sample selection of 5,875 M dwarfs after applying selection criteria described in Section \ref{subsec:test_selection} (right), to reduce contamination from sources with that are not similar to the training sample (not single, main-sequence M stars, such as pre-main sequence, spectroscopic binaries, and K dwarfs). Overplotted with orange triangles are the M15 and W11 training samples, for their respective \cannon\ test labels. Temperature gradient increases with decreasing color, spectral subtype increases with increasing color, and metallicity gradient increases perpendicularly up from the main sequence branch as expected. Deviations from these gradients seen at the upper slice of the main sequence is likely remaining contamination from the binary sequence.}
	\label{fig:safe_selection}
\end{center}
\end{figure*}

>>>>

\newcommand{\code}[1]{\texttt{\detokenize{#1}}}
\newcommand{\apogee}{\textsl{APOGEE}}
\newcommand{\thecannon}{\textsl{The Cannon}}
\newcommand{\cannon}{\textsl{Cannon}}
\newcommand{\gaia}{\textsl{Gaia}}
\newcommand{\drtwo}{\textsl{Gaia-DR2}}
\newcommand{\zmass}{\textsl{2MASS}}
\newcommand{\sdss}{\textsl{SDSS}}
\newcommand{\kepler}{\textsl{Kepler}}
\newcommand{\lamost}{\textsl{LAMOST}}
\newcommand{\sdssiv}{\textsl{SDSS-IV}}
\newcommand{\tess}{\textsl{TESS}}
\newcommand{\msun}{M_{\odot}}
\newcommand{\teff}{T_{\mathrm{eff}}}
\newcommand{\logg}{\log g}
\newcommand{\feh}{[{\mathrm{Fe}/\mathrm{H}}]}
\newcommand{\mh}{[{\mathrm{M}/\mathrm{H}}]}

@@

\item \textbf{Model extraploation cuts:} Because \thecannon\ does not perform well under extrapolated regions of parameter space, we select only sources inside the range of our training sample with ASPCAP parameters $2800<T_{\rm eff}<4100$ and $-0.5<\mh<0.5$, and with \cannon\ parameters of $2850<T_{\rm eff}<4150$, $-0.5<\feh<0.5$, and $0<SPT<9$.

@@

\begin{deluxetable*}{cccccccc}
\tablecaption{Metallicities of \apogee\ M dwarfs with wide binaries. Reported in this table are the ASPCAP [M/H] and \cannon\ [Fe/H] of the M dwarfs, as well as the metallicity values of their binary companions. The source/survey of the metallicity and spectral classes of the companions are also given. \label{table:binaries}}
\tabletypesize{\scriptsize}
\tablehead{Mdwarf Gaia ID      & Companion Gaia ID   & Mdwarf [M/H] & Mdwarf [Fe/H] & Companion [Fe/H] & Companion source & Separation (AU)}
	\startdata
	1476768433933264128 & 1476762627137479552 & -0.18      & 0.01        & 0.10           & APOGEE Cannon M   & 16091      \\
	1476646285062821888 & 1476646250703083264 & -0.02      & 0.18        & 0.27           & APOGEE Cannon M   & 4155       \\
	4008743154907642496 & 4008743086188165248 & -0.20      & 0.02        & 0.09           & APOGEE Cannon M   & 11855      \\
	3053911383155592832 & 3053907221325174144 & -0.25      & -0.15       & -0.15          & APOGEE Cannon M   & 12884      \\
	1475867727751525632 & 1475867727751525504 & -0.25      & -0.15       & -0.13          & APOGEE Cannon M   & 5091       \\
	933708229245319552  & 933708164821511424  & -0.23      & 0.11        & 0.01           & APOGEE Cannon M   & 2245       \\
	3888520938618496768 & 3888520938618115456 & -0.16      & -0.15       & -0.14          & APOGEE Cannon M   & 1128       \\
	3888309828090932608 & 3888309832385988992 & -0.24      & -0.01       & 0.04           & APOGEE Cannon M   & 2499       \\
	374137560388179328  & 374137560388178304  & -0.34      & -0.24       & -0.34          & LAMOST FGK        & 5888       \\
	4031346880591589248 & 4031346914951327104 & -0.33      & -0.25       & -0.18          & LAMOST FGK        & 6637       \\
	3443427533502334592 & 3443427533502335744 & 0.06       & 0.20        & 0.32           & LAMOST FGK        & 3151       \\
	1605991802162861696 & 1605991767803123200 & -0.16      & -0.03       & -0.06          & LAMOST FGK        & 1043       \\
	4418850890305439872 & 4418851263967008000 & -0.18      & 0.11        & 0.27           & APOGEE ASPCAP FGK & 4920       \\
	1381455722290898176 & 1381408782592655616 & -0.13      & -0.02       & -0.08          & APOGEE ASPCAP FGK & 2401       \\
	2687603208839189376 & 2687603311918406528 & -0.12      & 0.00        & -0.03          & APOGEE ASPCAP FGK & 11201    
	\enddata
\end{deluxetable*}

@@

\begin{figure}
\begin{center}
	\includegraphics[width=\linewidth]{figures/wide_binary_met.png}
	\caption{Metallicities of \cannon\ M dwarfs in binaries (x-axis) compared to the metallicities of their companions (y-axis). Label colors show the source and spectral type of the companion metallicities. Metallicity values for each binary pair is compiled in Table \ref{table:binaries}. }
\end{center}
\end{figure}

@@

As an additional check of our metallicity consistency, we cross-matched our M dwarf final sample with the \citealt{Elbadry:2018b} catalog of $>$50,000 high-confidence, widely separated binaries identified by \drtwo. 
In total we found 230 of the \apogee\ M dwarfs to have binary pairs (XX FGK+M, XX M+M, and XX WD+M).

As expected, the Toomre diagram in Figure \ref{fig:toomre} shows that higher metallicity sources in the sample are concentrated in low velocity space corresponding roughly to the thin disk population, while the thick disk population contains a higher concentration of lower metallicity sources. Separating the two populations into separate histograms (also shown in Figure \ref{fig:toomre}), we find that thick disk sources are nearly 1-$\sigma$ more metal poor than thin disk sources with the average thin disk metallicity being near-solar $\feh=0.00\pm0.17\,$dex, and the average metallicities of the thick disk stars being $\feh=-0.14\pm0.19\,$dex.
By contrast metallicities from ASPCAP show no significant difference in the distribution with an average thin disk metallicity of $\mh=-0.16\pm0.16\,$dex and an average thick disk metallicity of $\mh=-0.23\pm0.17\,$dex.

@@

\begin{figure*}
\begin{center}
	\includegraphics[width=.45\linewidth]{figures/cmd_teff_full.png}
	\includegraphics[width=.48\linewidth]{figures/cmd_teff_safe.png} \\
	\includegraphics[width=.45\linewidth]{figures/cmd_feh_full.png}
	\includegraphics[width=.48\linewidth]{figures/cmd_feh_safe.png} \\
	\includegraphics[width=.45\linewidth]{figures/cmd_spt_full.png}
	\includegraphics[width=.48\linewidth]{figures/cmd_spt_safe.png}
	\caption{Full sample of 14,827 M dwarfs colored by \cannon\ labels before selection (left) and final sample selection of 5,875 M dwarfs after applying selection criteria described in Section \ref{subsec:test_selection} (right), to reduce contamination from sources with that are not similar to the training sample (not single, main-sequence M stars, such as pre-main sequence, spectroscopic binaries, and K dwarfs). Overplotted with orange triangles are the M15 and W11 training samples, for their respective \cannon\ test labels. Temperature gradient increases with decreasing color, spectral subtype increases with increasing color, and metallicity gradient increases perpendicularly up from the main sequence branch as expected. Deviations from these gradients seen at the upper slice of the main sequence is likely remaining contamination from the binary sequence.}
	\label{fig:safe_selection}
\end{center}
\end{figure*}
